---
title: "The Effect of Vitamin C on Tooth Growth in Guinea Pigs"
author: "JRB"
date: "September 16, 2016"
output: pdf_document
geometry: margin=1in
fontsize: 8pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr);library(knitr);library(pander)
require(broom);require(stargazer)
panderOptions('digits',2)
panderOptions('round',2)
```

##Overview
We propose to analyze the length of odontoblasts for 60 guinea pigs that have received one of three dose levels of Vitamin C by of of two delivery methods (orange juice OJ) or ascorbic acid (VC). We are trying to assert if Vitamin C effect the tooth growth of guinea pigs and if it does, does the mean of administration impact the growth.   

##Exploratory Data Analysis 

We fill first plot the data set to represent the density of observations and some key statistics for each combination of delivery method at each dose level. The graph will compare side by side the tooth length for dose level for both OJ and VC. 


```{r load dataset and some basic plotting, echo=FALSE,fig.align='center',fig.height=4, fig.width=6}
data("ToothGrowth")

library(ggplot2)
g <- ggplot(
                data=ToothGrowth,
                aes(x=interaction(supp,dose),y=len)
                )
g + geom_violin(
                aes(fill=supp),
                scale="area") + 
    stat_summary(fun.data ="mean_se",
                 colour="white") + 
    scale_fill_brewer(palette="Blues") + 
    theme_minimal() + 
    geom_boxplot(width=0.1) +
    ggtitle("Tooth length by dose and supplement") +
    xlab("Supplement and Dose") +
    ylab("Length")
```
The figure illustrates that increasing doses of vitamin C would correspond to 
longer teeth and that the mean of admistration could impact tooth growth also.
Table 1 provides a summary of the `ToothGrowth` data set.
```{r stats, echo=FALSE}
        tbTooth <- tbl_df(ToothGrowth)
        ToothSummaryAll <- tbTooth %>% 
                        group_by(dose,supp) %>% 
                        summarise( 
                                n(),
                                mn =mean(len),
                                s=round(sd(len),2),
                                sem=round(sd(len)/sqrt(n()),2)
                                )
        ToothSummaryDose <- tbTooth %>% 
                        group_by(dose) %>% 
                        summarise( 
                                n(),
                                mn =mean(len),
                                s=round(sd(len),2),
                                sem=round(sd(len)/sqrt(n()),2)
                                )
    
```
  
  
```{r table1 }
            ToothSummarySupp <- tbTooth %>% 
                        group_by(supp) %>% 
                        summarise( 
                                n(),
                                mn =mean(len),
                                s=round(sd(len),2),
                                sem=round(sd(len)/sqrt(n()),2)
                                )
        names(ToothSummarySupp)[-1] <- c("n","Mean","Std.Dev","Sample Error")
```


From the summary table we can formulate the following hypothesis that could be submitted to a statistical testing.
The hypothesis are:  

#. Tooth length increases as dose increases
#. Mean of delivery has an impact on tooth growth 

```{r table3}
        names(ToothSummaryAll)[-(1:2)] <- c("n","Mean","Std.Dev","Sample Error")
        pander(ToothSummaryAll,caption="Summary Statistics for tooth length by dose and supplement",alignment=c("center"))
```        

##Hypothesis Testing  

###Hypothesis 1:  "Tooth length increases as dose increases"  
To test this hypothesis, we will compare tooth length means at different dose levels (regardless of mean of administration). We will test the null hypothesis that the samples are drawn from populations of the same mean. 

####*Example: Test 1: 0.5 mg vs 1 mg* 
This section explains our general methodology for hypothesis testing. All results will be summarized in table 2.
$\mu_1$ is the mean of tooth length for the 1 mg group  
$\mu_2$ is the mean of tooth length for the 0.5 mg group  
$H_0: \mu_1=\mu_2$, $H_3: \mu_1>\mu_2$    
```{r confidence interval for dosewise comparison,include=FALSE}

D0.5 <- subset(ToothGrowth,dose==0.5)
D1 <- subset(ToothGrowth,dose==1)
D2 <- subset(ToothGrowth,dose==2)
D105 <- subset(ToothGrowth,dose %in% c(1,0.5))
D12 <- subset(ToothGrowth,dose %in% c(2,1))
D205 <- subset(ToothGrowth,dose %in% c(2,0.5))

ttest <- NULL
ttest <- cbind(t.test(D1$len,D0.5$len,alternative="greater"),
               t.test(D2$len,D1$len,alternative="greater"),
               t.test(D2$len,D0.5$len,alternative ="greater")
               )
#formatting the data name

colnames(ttest) <- c("Test 1: 0.5 vs 1 mg","Test 2: 1 vs 2 mg", "Test 3: 0.5 vs 2 mg")
rbind(ttest,c(
        round(qt(0.95,ttest[,1]$parameter),2),
        round(qt(0.95,ttest[,2]$parameter),2),
        round(qt(0.95,ttest[,3]$parameter),2)
      ))

```
First we will calculate our t-statistics (TS) using the R `t.test` function. For our statistical test we assume an unequal variance and a single sided test. 
We will reject $H_0$ if $TS \geqslant t_{0.95}$, the $t_{0.95}$ is calculated using `r round(t.test(D1$len,D0.5$len,type="greater")$parameter,2)` as the degree of freedom for unequal variances and provided by the `t.test` function. $t_{0.95}=$ `r round(qt(0.95,round(t.test(D0.5$len,D1$len,type="greater")$parameter,2)),2)`. TS is also provided by the R `t.test` function: TS= `r round(t.test(D1$len,D0.5$len,type="greater")$statistic,2)`. We do have $TS \geqslant t_{0.95}$, and we therefore reject our null hypothesis in favor of the alternate. We conlcude that tooth length is on average greater for the 1mg dose than for the 0.5 mg dose with a 95% confidence. 
`r pander(ttest[c(1,2,3),],digit = 1) `

As the previous table show, in all cases the Test Statistics is over the t-value for 0.95 ($t_{0.95}=1.69$) and the appropriate degree of freedom (df~38). We can appropriately reject $H_0$. There's a difference in the mean and higer doses have a higher mean than lower doses.

###Hypothesis 2: Mean of delivery has an impact on tooth growth   

####*Impact of mean of administration on tooth growth across all dose groups*  
$\mu_1$ is the mean of tooth length for the OJ group  
$\mu_2$ is the mean of tooth length for the VC group  

$H_0: \mu_1=\mu_2$, $H_3: \mu_1>\mu_2$  
```{r t.test len ~ supp, echo=FALSE}
        testlenbysupp <- t.test(data=ToothGrowth,len ~ supp,alternative="greater")
```
`r pander(testlenbysupp)`  

The t-value ($t_{0.95}$) with `r round(testlenbysupp$parameter,2)` degree of freedom is `r round(qt(0.95,testlenbysupp$parameter),2)`. The $t_{0.95}$ value is less than the test statistics so we can reject the null hypothesis.  
***Conclusion*** When analyzed across dose groups the mean of administration does have an impact on tooth growth. We will now refine our analysis and look at the tooth length mean differences at each dose level comparing the two means of administration.  

####*Mean of admistration impact by dose level*  
$\mu_1$ is the mean of tooth length for the OJ group at the analysed dose level  
$\mu_2$ is the mean of tooth length for the VC group at the analysed dose level 

$H_0: \mu_1=\mu_2$, $H_3: \mu_1>\mu_2$    

```{r t.test len ~supp @ 0.5mg, echo=FALSE}
        
        res<- cbind(t.test(data=D0.5,len ~ supp,alternative="greater"),
                    t.test(data=D1,len ~ supp,alternative="greater"),
                    t.test(data=D2,len ~ supp,alternative="greater")
                        )
        colnames(res) <- c("len vs. supp at 0.5 mg","len vs. supp at 1 mg","len vs. supp at 2 mg")
  res0.5 <- res[,1]                  
```
```{r, results="asis"}
pander(res[c(1,2,3),])
```
  

At the 0.5 mg and 1 mg dose level, we accept the alternate hypothesis, $\mu_1$ > $\mu_2$ ($p<\alpha$). At the 2 mg level we do not reject the null hypothesis.  
***Conclusion*** At the 0.5mg and 1mg dose level, OJ has a greater mean tooth length than VC. At this dose level the mode of administration has an impact on tooth growth. At the 2 mg level the mean tooth length for OJ and VC is the same.  

##Conclusion and notes on assumptions
During our analysis we assumed that the samples were iid, the sample variances were considered unequal and we used a confidence level of 95% for all statistical testing. 
With the statistical rigor of our hypothesis testing we can conclude that:  

#. As the dose of vitamin C increases so does the length of the tooth. Vitamin C has a positive impact on the tooth growth of guinea pig  
#. The mean of admnistration Orange Juice or Ascorbic Acid has an impact on two growth overall. The impact is measurable for the lower dose groups (0.5mg and 1mg) and favors orange juice over ascorbic acid  

\newpage
#Appendix
## Additional Summary Tables 


```{r table 1}
pander(ToothSummarySupp,caption ="Summary statistics for tooth length by supplement",alignment="center") 
```
  


```{r table2}
        names(ToothSummaryDose)[-1] <- c("n","Mean","Std.Dev","Sample Error")
        pander(ToothSummaryDose,caption="Summary statistics for tooth length by dose",alignment="center")
```
